---
title: Interest Payments plot with ggplot package
author: DF
date: '2018-11-18'
slug: interest-payments-plot-with-ggplot-package
categories:
  - economics
tags:
  - R
  - ggplot
  - plot
---

  A lot of things have been said and written about greek statistics and greek economy through the last years.

Unfortunately, there are not many sources of data about key indicators of greek economy and that has contributed in a poorly informed public dialogue. Though, Bank of Greece has uploaded a few datasets in a dedicated website *(https://opendata.bankofgreece.gr/en/home)* so we can use them with *R* to shed some light in greek economic reality.

We can start with a very simple and straight forward plot that depicts the interest payments for the greek public debt from 2010 to 2017.

# Preparing the Data

First we have  to load a few libraries to our workspace. We can do that either with library() or with require() function.

```{r, warning=FALSE,message=FALSE}
require(data.table) ; require(installr)
require(dplyr) : require(ggplot2) ; require(lubridate)
```

Bank of Greece provides data only in the .xlsx format so we will download the file automatically and convert it through R with xlsx2csvv() function.

```{r, warning=FALSE, message=FALSE}
dataurl <- "https://opendata.bankofgreece.gr/datasetFile.ashx?fileName=BoG_CG_NBR_Cash_en_2018-06-26.xlsx&folderName=DOAM"

download.file(dataurl, destfile = "BoG_CG_NBR_Cash_en_2018-06-26.xlsx", mode = "wb") 

# now we will convert the file and we will store the data in a dataframe

if(file.exists("BoG_CG_NBR_Cash_en_2018-06-26.xlsx")) {
  
  xlsx2csv("BoG_CG_NBR_Cash_en_2018-06-26.xlsx", "BoG_CG_NBR_Cash_en_2018-06-26.csv") # convert the xlsx to csv
  
  df <- fread("BoG_CG_NBR_Cash_en_2018-06-26.csv", header = T)
}
```
# Clear the Dataframe
```{r, warning=FALSE,message=FALSE}
# delete the empty variables

df$V7 <- NULL
df$V8 <- NULL
df$V9 <- NULL

# change date variable to date format with lubridate package of the tidyverse

df$From <- ymd(df$From)
df$To <- ymd(df$To)

df$year <- year(df$To)  



```
# Making the Plot using the ggplot package

```{r, warning=FALSE, message=FALSE}

# first make the basic plot 
plot_hist_interest <- ggplot(df, aes(x=To, y = Interest_payments)) 

# then add the aesthetics

plot_hist_interest + geom_col(alpha = 2/10 ,color = "steelblue") +
  labs(title = "Interest Payments by month (cumulative)", subtitle = "Greek Economy",
       x =  "Year",
       y = "Interest Payments (In million euros)") +
  theme(plot.title = element_text(hjust = 0.5) ) +
  theme(plot.subtitle = element_text(hjust = 0.5))

```


From the above plot it is more than obvious that after 2012 the interest payments are substuntially reduced. This reduction was a direct result of a debt rectucture known as Private Sector Involvement. 

Greece accomplished through PSI a very generous haircut, almost 50%, of its bonds, through a highly complex procedure. If you are interested in the legal aspect of the PSI which had made use of advanced and new legal tools you may read a very thorough paper published in CESifo Working Paper Series (https://scholarship.law.duke.edu/cgi/viewcontent.cgi?article=5343&context=faculty_scholarship)




